Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses

### 2018-04-16 paper-webserver.tex
EDIT: fig positions
START: mongo + total runtimes

Spfy provides rapid \textit{in-silico} versions of common reference laboratory tests for the analyses of \textit{E. coli}. It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}; and AMR annotation using the resistance gene identifier from the comprehensive antibiotic resistance database \cite{mcarthur2013comprehensive}. Spfy also performs pan-genome analyses using Panseq \cite{laing2010pan}, with the entire pan-genome stored in the database and associated with source genomes.

@@

@@

@@

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/results.png}
\end{center}
\caption{The results interface for submitted tasks. Cards represent individual tasks and the entire collection can be accessed by sharing the URL with the embedded token.}
\label{fig-results}
\end{figure}

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/tables.png}
\end{center}
\caption{Detailed results for the virulence factor subtyping task. While data storage in Spfy is graph-based, a familiar tabular structure is presented to users. The result can also be shared using the URL with the embedded token.}
\label{fig-tables}
\end{figure}

@@

On comparison tasks, Spfy can retrieve and compare 1 million nodes or attributes in the graph database in aproximately 70 seconds \ref{fig:fishers_performance}.
Performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome.

@@

On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance}.
Spfy can also queue up batches of files for analysis, decreasing the average runtime to 54 seconds on average with a degree of parallelization.

>>>>

Spfy provides rapid \textit{in-silico} versions of common reference laboratory tests for the analyses of \textit{E. coli}. It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}; and AMR annotation using the resistance gene identifier from the comprehensive antibiotic resistance database \cite{mcarthur2013comprehensive}. An example of the VF results is given in \ref{fig-tables}. Spfy also performs pan-genome analyses using Panseq \cite{laing2010pan}, with the entire pan-genome stored in the database and associated with source genomes.

\begin{figure}[!htb]
\begin{center}{}
\includegraphics[width=\textwidth]{images/tables.png}
\end{center}
\caption{Detailed results for the virulence factor subtyping task. While data storage in Spfy is graph-based, a familiar tabular structure is presented to users. The result can also be shared using the URL with the embedded token.}
\label{fig-tables}
\end{figure}

@@

Spfy primarily uses Blazegraph \url{https://github.com/blazegraph/database} for storage along with MongoDB to cache a hash table for duplicate checking.
The cache allows Spfy to more efficiently check for duplicate files in Blazegraph, than would be possible through an exaustive search of the graph structure.

@@

\begin{figure}[!htb]
\begin{center}
\includegraphics[width=\textwidth]{images/results.png}
\end{center}
\caption{The results interface for submitted tasks. Cards represent individual tasks and the entire collection can be accessed by sharing the URL with the embedded token.}
\label{fig-results}
\end{figure}

@@

@@

Spfy is running on a virtual machine (VM) with 8 vCPUs emulating single-core Intel® Xeon® E3-series processors, and with 32 GB of RAM.
The VM is running Linux with CentOS.
On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in aproximately 70 seconds.
As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in aproximately 90 seconds, and 2 million nodes/attributes in aproximately 110 seconds.

@@

On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance}.
Spfy can also queue up batches of files for analysis, decreasing the average runtime to 54 seconds on average with a degree of parallelization.
% TODO: add totals
